A New Metropolis-Hastings Test for Fast Minibatch MCMC Inference and Optimization

### 2016-05-15 nips_2016.tex
major update to sections 1 through 4. Obviously not done but lots of progress in a draft form.

\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem}
\graphicspath{{figures/}}

@@

Therefore, we are not interested in exact inference of the posterior model parameters for a very
large dataset, but rather, we want to perform either MAP or ML estimation. In both cases the goal is
to find a good posterior mode, but the exact form of the distribution is arbitrary. In particular,
non-unit powers of the distribution are fine. That corresponds to changing the temperature of the
distribution to make it ``flatter''.

Rather than fixing the posterior temperature and trying to use adjustable mini-batch size to achieve
it (which ends up reducing to nearly-full-batch updates) we propose to run at the natural
temperature induced by a given minibatch size. We can always increase the minibatch size to reduce
the temperature, but there is no need to do this beyond the point where the posterior mode is sharp
and well-defined. That will occur at some fixed number of points and far less than the full dataset.

To this end, we present a new Metropolis-Hastings test that can fit in this framework of minibatch
MCMC methods. We present theoretical and experimental results on our method and show that it can be
the method of choice for minibatch MCMC .

@@

The standard MCMC method proceeds as follows~\cite{gilks1996markov,brooks2011handbook}. For a
parameter $\theta$ and conditionally independent data, we wish to compute the target distribution
$p(\theta \mid x_1, \ldots, x_N)$. Since exact evaluation is intractable, we generate a chain of
correlated samples $\theta_1, \ldots, \theta_T$ for some large $T$, and approximate $p$ by using
counts of samples. For each iteration $t$, we start with our current $\theta_t$. We use a
\emph{proposal distribution} $q(\theta' \mid \theta_t)$ to determine a new candidate $\theta'$. With
probability $P_a$, we accept it and set $\theta_{t+1} = \theta'$; otherwise, we repeat the previous
value $\theta_{t+1} = \theta_t$. Traditionally, $P_a$ is computed as follows:
\begin{equation}\label{eq:traditional}
P_a = \min\left\{ 1, \frac{f(\theta')q(\theta_t \mid \theta')}{f(\theta_t)q(\theta' \mid \theta_t)}
\right\} = \min\left\{ 1, \frac{p(\theta')\prod_{i=1}^N p(x_i ; \theta')q(\theta_t \mid \theta')}{p(\theta_t)\prod_{i=1}^N p(x_i ; \theta_t)q(\theta' \mid
\theta_t)} \right\},
\end{equation}
where $f(\theta_t)=p(\theta_t|x_1,\ldots,x_N) \propto p(\theta_t)\prod_{i=1}^N p(x_i \mid
\theta_t)$. The technical requirement for $P_a$ is that $f$ be proportional to the true target
distribution.

One then draws a uniform random variable $u$ from ${\rm Unif}[0,1]$ and accepts $\theta_{t+1} =
\theta'$ if $u < P_a$, and uses $\theta_{t+1} = \theta_t$ otherwise. The $P_a$ from
Equation~\ref{eq:traditional} satisfies detailed balance, so if one samples long enough using the
heuristic above, one will arrive at a stationary distribution.  The $P_a$ is guaranteed to converge
to the true $p$ distribution given sufficiently many examples (though a burn-in period and/or taking
every $n$th sample may be necessary).

@@

{\color{blue}
Daniel: I am worried. We need to motivate why it is not easy to use minibatch MCMC, and what other
researchers have tried to do about that.  The last paragraph is from~\cite{cutting_mh_2014}, and it
seems to argue that people have not used minibatch MCMC before ``the algorithm gets stuck''. But
isn't that exactly what our algorithm is doing when it is getting ``stuck'' at modes? If we wanted
to get stuck at modes with large datasets, why didn't we just do what previous work did?
}

The work of~\cite{cutting_mh_2014} proposes an adaptive minibatch approach by using a sequential
hypothesis testing algorithm. During each iteration, they start with a small minibatch of data and
test the hypothesis that the sample $\theta'$ should be accepted or rejected based the likelihood
ratio. If the test cannot make a decision over a certain confidence threshold, then they increase
the minibatch size and test again. This process repeats until a decision.  The downside of this
algorithm is the need to keep incrementing the mini-batch size in one iteration; in the worst case,
all $N$ data points may be needed in a single iteration. A related approach
is~\cite{icml2014c1_bardenet14}, which has slightly more robust theoretical properties, but comes at
the cost of more computational time. The work of~\cite{conf/uai/MaclaurinA14} presents an
\emph{exact} approach, but has the sometimes unreasonable requirement of a lower bound on the
per-datum likelihood factors.

@@

Our starting point is the following Lemma.

\begin{lemma}\label{lem:detailed_balance}
Let $\Delta = \log \left(\frac{f(\theta') q(\theta_t \mid \theta')}{f(\theta_t) q(\theta'\mid
\theta_t)} \right)$, where $f$ is proportional to the desired target distribution and $q$ is our
chosen proposal distribution. Any acceptance function $g$ such that $g(\Delta) = \exp(\Delta)
g(-\Delta )$ satisfies detailed balance. That is, $f(\theta_t)p(\theta' \mid \theta_t) =
f(\theta')p(\theta_t \mid \theta')$, where $p(\theta_y \mid \theta_x)$ is the probability of jumping
from $\theta_x$ to $\theta_y$ in our chain.
\end{lemma}

\begin{proof}
We begin by deriving $p(\theta' \mid \theta_t)$. This is equivalent to the probability of proposing
$\theta'$ and then accepting it, so
\begin{equation}\label{eq:one_way}
p(\theta' \mid \theta_t) = q(\theta' \mid \theta_t)g(\Delta).
\end{equation}
Similarly, we have
\begin{equation}\label{eq:other_way}
p(\theta_t \mid \theta') = q(\theta_t \mid \theta')g(-\Delta).
\end{equation}
Notice that the probability of accepting a transition from $\theta'$ to $\theta_t$ is $g(-\Delta)$
because this inverts the fraction inside the logarithm term of $\Delta$.  By assumption, we can
expand $g(\Delta) = \exp(\Delta)g(-\Delta)$ in Equation~\ref{eq:one_way}.  Doing this, and combining
the result of Equation~\ref{eq:other_way}, we get
\begin{equation}\label{eq:combined}
g(-\Delta) = \frac{p(\theta' \mid \theta_t)}{q(\theta' \mid \theta_t)\exp(\Delta)} = \frac{p(\theta_t \mid \theta')}{q(\theta_t \mid \theta')}.
\end{equation}
Rearranging terms and expanding $\exp(\Delta)$, we have
\begin{equation}\label{eq:rearrange}
\frac{p(\theta' \mid \theta_t) f(\theta_t) q(\theta' \mid \theta_t)}{q(\theta' \mid \theta_t) f(\theta') q(\theta_t \mid \theta')} = \frac{p(\theta_t \mid \theta')}{ q(\theta_t \mid \theta')}.
\end{equation}
Cancellations result in $f(\theta') p(\theta_t \mid \theta') = f(\theta_t) p(\theta' \mid \theta_t)$. Thus, detailed balance is satisfied.\\
\end{proof}

@@

For reasons that will be clear later, we choose an alternative $g$, the logistic function:
$g(\Delta) = (1+\exp(-\Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the
condition in Lemma~\ref{lem:detailed_balance}.

We can sample from the logistic function using the following procedure. Let $u$ be drawn from ${\rm
Unif}[0,1]$. At any given iteration, we can compute $\Delta$, and we accept the new candidate
$\theta'$ if $g(\Delta) > u$, and reject otherwise. Let us define the random variable $X =
g^{-1}(u)$ where again, $g$ is the logistic function. We claim that $X$ has CDF function $F_X(x) =
g(x)$, i.e., its CDF is precisely the logistic function. This is because for some $X = x$, we have
\[
F_X(x) = {\rm Pr}(X \le x) = {\rm Pr}(g^{-1}(u) \le x) = {\rm Pr}(u \le g(x)) = \int_{0}^{g(x)} 1 dx = g(x),
\]
as the density of the uniform random variable here is simply one. Thus, the criteria to accept the
candidate $\theta'$ is equivalent to whether $\Delta > X$, where $X$ has CDF of logistic function.
(If this isn't immediately obvious, note that we can pre-multiply $g^{-1}$ to $g(\Delta) \ge u$, and
our test is $\Delta > X$.) Since $X$ is symmetric about zero, the acceptance criteria can also be
expressed as $\Delta + X>0$, as shown in Figure~\ref{fig:part1}.

{\color{blue}
Daniel: The above makes sense, but I still don't understand why we had to show that the CDF of $X$
is the logistic function, because I never use that when I claim that our acceptance test is now
$\Delta > X$.  Do we need that assumption so that, for instance, we can decompose $X$ later the way
we do it, with $X_{\rm noise}$ having the same variance as $\epsilon$? That might explain it.
}

Unfortunately, there's a problem with the $\Delta + X > 0$ test: it requires us to compute $\Delta$,
so our test will be slow! We can get around this by defining a new scalar-valued term, $\Delta'$, which is
computed the same way as $\Delta$, but uses far fewer samples. In other words, $\Delta'$ is computed
based on a \emph{mini-batch} of data, and it is an approximation of $\Delta$, so we can express the
relationship as $\Delta' = \Delta + \epsilon$ for a noise term $\epsilon$. \emph{The key is that
$\epsilon$ follows a normal distribution}, $\epsilon \sim \mathcal{N}(0, \sigma^2)$, because for i.i.d. data $x_1, \ldots, x_N$, the $\Delta$
is expressed as the sum of log probabilities (look at Equation~\ref{eq:traditional} and apply the
logarithm to the product over the $N$ data points, to get a sum of log terms). All we do with
$\Delta$ is take far fewer of the summations, and normally-distributed noise is there as a product
of the Central Limit Theorem.

To sample accurately from $\Delta'$, we need to slightly change our acceptance criteria. We can
decompose $X$ as $X = X_{\rm norm} +X_{\rm correction}$, where the first term is a ``noise term''
which has the same variance as $\epsilon$, and the second term is a ``residual'' term. The criteria
to accept, previously expressed as $\Delta + X > 0$, can be rewritten as
\begin{equation}\label{eq:criteria}
\Delta + X = \Delta + X_{\rm norm} + X_{\rm correction} \approx \Delta' + X_{\rm correction} >0.
\end{equation}
This explains Figure~\ref{fig:part2} (though it really should be an approximation as we can't guarantee that our $X_{\rm norm}$ term is exactly the correct amount of noise), where the left hand side has distributions representing normal
noise ($X_{\rm norm}$, which is labeled as ${\rm Noise}_U$, but they are the same thing) and a
``correction'' term labeled $X_{\rm correction}$. Be aware that all of these ``additions'' are to be
interpreted as convolutions when computing the sum of densities.

Notice that we have several options we can control. For instance, we can adjust the mini-batch size;
increasing the mini-batch size will cause the $X_{\rm noise}$ distribution to shrink and become
narrower.

{\color{blue} Daniel: I think it's becoming clear but I'm still trying to connect all of the steps.
During each iteration of MCMC, we compute $\Delta'$ and \emph{sample} (right?) an $X$ (which we can
do by sampling $u$ and then doing $X = g^{-1}(u)$. This gives us $\Delta' + X$ but our MH test
    actually requires $X_{\rm correction}$, so first we must estimate $\sigma^2$, the variance of
    $X_{\rm noise} \sim \mathcal{N}(0, \sigma^2)$ (perhaps from the mini-batch data, somehow?) and
    then somehow ``remove'' that component from $X$ to get $X_{\rm correction}$? Then we have our
    acceptance test. Is this right? Or, of course, we don't have to sample $X$ at all, but just
    figure out a way to sample $X_{\rm correction}$ directly.

I think we talked about estimating $\sigma^2$. The variance $\sigma^2$ that we're concerned about is
a scalar value based on the sum of log probabilities of different samples, so one way we could
estimate the variance is by taking all samples of a mini-batch and seeing how much those individual
$\log p(x_i \mid \theta)$ vary. Or we could look at $k$ mini-batches of data which will give us
$\Delta_1, \ldots, \Delta_k$ and compute the empirical $\sigma^2$ (of course that defeats the
purpose of using mini-batches).  }

{\color{blue} Xinlei: I agree with you that out MH test requires $X_{correction}$. My understanding
is that we can actually estimate $X$ from the whole data set, and the probability distribution of
$X_{norm}$ is what we proposed? So the determination of $X_{correction}$ is a deterministic
deconvolution problem.  }

@@

TODO insert Haoyu's proof.

>>>>

\usepackage{amsmath,amsthm,color,graphicx,verbatim,listings,enumitem}
\usepackage[]{algorithm2e}
\graphicspath{{figures/}}

@@

Therefore, we are not interested in exact inference of the posterior model parameters for a very
large dataset, but rather, we want to perform either MAP or ML estimation. In both cases the goal is
to find a good posterior mode, but the exact form of the distribution is arbitrary. In particular,
non-unit powers of the distribution are fine. That corresponds to changing the temperature of the
distribution to make it ``flatter''.  Rather than fixing the posterior temperature and trying to use
adjustable minibatch size to achieve it (which ends up reducing to nearly-full-batch updates) we
propose to run at the natural temperature induced by a given minibatch size. We can always increase
the minibatch size to reduce the temperature, but there is no need to do this beyond the point where
the posterior mode is sharp and well-defined. That will occur at some fixed number of points and far
less than the full dataset.

To this end, we present a new Metropolis-Hastings test that can fit in this framework of minibatch
MCMC methods. We present theoretical and experimental results on our method and show that it can be
the method of choice for minibatch MCMC. The contributions of this paper are as follows:

\begin{enumerate}[noitemsep]
    \item We develop a new minibatch MH test which satisfies detailed balance.
    \item Our MH test is more efficient than the adaptive sampling method.
    \item The test can incorporate temperature considerations to control mixing rates.
\end{enumerate}

{\color{blue}
Daniel: This section needs work in describing our argument carefully. I will decrease the focus on
temperature and increase focus on the new MH test once the results are clearer.
}

@@

In the standard MCMC method~\cite{gilks1996markov,brooks2011handbook} for Bayesian inference with
parameter $\theta$ and conditionally independent data, the goal is to compute the target
distribution $p(\theta \mid x_1, \ldots, x_N)$.  To do so, one generates a chain of (correlated)
samples $\theta_1, \ldots, \theta_T$ for large $T$ and approximates $p$ using the sample counts.
Each iteration $t$ has a current $\theta_t$, and a \emph{proposal distribution} $q(\theta' \mid
\theta_t)$ determines a new candidate $\theta'$. With probability $P_a$, the sample is accepted, so
$\theta_{t+1} = \theta'$. Otherwise, $\theta_{t+1} = \theta_t$. This can be implemented by drawing
a uniform random variable $u \sim {\rm Unif}[0,1]$ and accepting if $u < P_a$. Traditionally, $P_a$ is
\begin{equation}\label{eq:traditional}
P_a = \min\left\{ 1, \frac{f(\theta')q(\theta_t \mid \theta')}{f(\theta_t)q(\theta' \mid \theta_t)}
\right\} = \min\left\{ 1, \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)q(\theta' \mid
\theta_t)} \right\},
\end{equation}
where $f(\theta_t)=p(\theta_t \mid x_1,\ldots,x_N)$. The $P_a$ from Equation~\ref{eq:traditional}
satisfies detailed balance, so if one samples long enough, one will arrive at a stationary
distribution matching the posterior, though a burn-in period and/or taking every $n$th sample may be
helpful.

@@

{\color{blue}
Daniel: I am worried. We need to motivate why it is not easy to use minibatch MCMC, and what other
researchers have tried to do about that.  The last paragraph is from~\cite{cutting_mh_2014}, and it
seems to argue that people have not used minibatch MCMC before because ``the algorithm gets stuck''.
But \emph{isn't that exactly what our algorithm is doing when it is getting ``stuck'' at modes}? If
we wanted to get stuck at modes with large datasets, why didn't we just do what previous work did?
The only way around this is if our algorithm is getting stuck at ``better'' modes, but then we need
to design our experiments to test this.
}

The work of~\cite{cutting_mh_2014} proposes an adaptive minibatch approach by using a sequential
hypothesis testing algorithm. During each iteration, they start with a small minibatch of data and
test the hypothesis that the sample $\theta'$ should be accepted or rejected based the likelihood
ratio. If the test cannot make a decision over a certain confidence threshold, then they increase
the minibatch size and test again. This process repeats until a decision.  The downside of this
algorithm is the need to keep incrementing the mini-batch size in one iteration; in the worst case,
all $N$ data points may be needed in a single iteration. A similar approach
from~\cite{icml2014c1_bardenet14} has slightly more robust theoretical properties, but comes at the
cost of more computational time. The work of~\cite{conf/uai/MaclaurinA14} presents another minibatch
MCMC method, but which gets the \emph{exact} posterior, with the only requirement being that there
needs to be a lower bound on the log likelihood for each data point. They augment the likelihood
with auxiliary random variables, but this can hurt mixing.

@@

In this section, we introduce our new Metropolis-Hastings test and show how to use it to achieve
sufficient mixing and/or to reach better posterior modes. To start, we introduce two key values we
will use frequently: $\Delta$ and $\Delta'$:
\begin{equation}\label{eq:deltas}
\Delta = \log \left( \frac{p(\theta')\prod_{i=1}^N p(x_i \mid \theta')q(\theta_t \mid
\theta')}{p(\theta_t)\prod_{i=1}^N p(x_i \mid \theta_t)q(\theta' \mid
\theta_t)} \right);\;  \Delta' = \log \left( \frac{p(\theta')\prod_{i=1}^n p(x_i \mid \theta')q(\theta_t \mid
\theta')}{p(\theta_t)\prod_{i=1}^n p(x_i \mid \theta_t)q(\theta' \mid
\theta_t)} \right)
\end{equation}
where $p$ and $q$ have similar definitions as in Equation~\ref{eq:traditional}, so the terms here
consider the likelihood ratios in log space. The only difference between $\Delta$ and $\Delta'$ is
that they use $N$ and $n$ terms in the likelihood ratio, respectively. We assume $n \ll N$ (so
$\Delta'$ is substantially faster to compute) and that the $n$ data points $x_i$ represent a random
minibatch of data.

To motivate the use of $\Delta$, we turn to Lemma~\ref{lem:detailed_balance}.

\begin{lemma}\label{lem:detailed_balance}
Let $\Delta$ be defined as in Equation~\ref{eq:deltas}. Any acceptance function $g$ such that
$g(\Delta) = \exp(\Delta) g(-\Delta )$ satisfies detailed balance. That is, $f(\theta_t)p(\theta'
\mid \theta_t) = f(\theta')p(\theta_t \mid \theta')$, where $p(\theta_y \mid \theta_x)$ is the
probability of jumping from $\theta_x$ to $\theta_y$ in our chain.
\end{lemma}

\begin{proof}
We begin by deriving $p(\theta' \mid \theta_t)$. This is equivalent to the probability of proposing
$\theta'$ and then accepting it, so $p(\theta' \mid \theta_t) = q(\theta' \mid \theta_t)g(\Delta).$
Similarly, $p(\theta_t \mid \theta') = q(\theta_t \mid \theta')g(-\Delta).$ Notice that the
probability of accepting a transition from $\theta'$ to $\theta_t$ is $g(-\Delta)$ because this
inverts the fraction inside the logarithm term of $\Delta$.  By assumption, we can expand $g(\Delta)
= \exp(\Delta)g(-\Delta)$ in $p(\theta' \mid \theta_t)$. Doing this, and combining that with the
definition of $p(\theta_t \mid \theta')$, we get
\begin{equation}\label{eq:combined}
g(-\Delta) = \frac{p(\theta' \mid \theta_t)}{q(\theta' \mid \theta_t)\exp(\Delta)} = \frac{p(\theta_t \mid \theta')}{q(\theta_t \mid \theta')}.
\end{equation}
Rearranging terms and expanding $\exp(\Delta)$, we have
\begin{equation}\label{eq:rearrange}
\frac{p(\theta' \mid \theta_t) f(\theta_t) q(\theta' \mid \theta_t)}{q(\theta' \mid \theta_t) f(\theta') q(\theta_t \mid \theta')} = \frac{p(\theta_t \mid \theta')}{ q(\theta_t \mid \theta')}.
\end{equation}
Cancellations result in $f(\theta') p(\theta_t \mid \theta') = f(\theta_t) p(\theta' \mid
\theta_t)$. Thus, detailed balance is satisfied.
\end{proof}

@@

For our MH test, the key is that we use a different $g$, the logistic function: $g(\Delta) =
(1+\exp(-\Delta))^{-1}$. Again, straightforward arithmetic shows that it satisfies the condition in
Lemma~\ref{lem:detailed_balance}.  The logistic function is nice because we can easily sample from
it using the following procedure. Let $u$ be a uniform random variable $u \sim {\rm Unif}[0,1]$. At
any iteration with a current parameter $\theta_t$ and a proposed sample $\theta'$, we can compute
$\Delta$. We accept the candidate if $g(\Delta) > u$, and reject otherwise. To see why this works,
define a new random variable $X = g^{-1}(u)$ where again, $g$ is the logistic function. Notice that
$X$ has cumulative distribution function $F_X(x) = g(x)$. In other words, the CDF is precisely the
logistic function. This is because for some $X = x$, we have
\[
F_X(x) = {\rm Pr}(X \le x) = {\rm Pr}(g^{-1}(u) \le x) = {\rm Pr}(u \le g(x)) = \int_{0}^{g(x)} 1 dx = g(x),
\]
as the density of the uniform random variable here is simply one. Thus, the criteria to accept the
candidate $\theta'$ is equivalent to whether $\Delta > X$. Since $X$ is symmetric about zero, the
acceptance criteria is equivalently expressed as $\Delta + X>0$, as shown in Figure~\ref{fig:part1}.

{\color{blue}
Daniel: The above makes sense, but I \emph{still} don't understand why we had to show that the CDF
of $X$ is the logistic function, because I never use that when I claim that our acceptance test is
now $\Delta > X$. All I need to say is this: we can pre-multiply $g^{-1}$ to $g(\Delta) \ge u$, and
our test is $\Delta > X$. Right???
}

With large datasets, $\Delta$ is intractable to compute. An intuitive alternative is to use
$\Delta'$ instead. The following lemma characterizes the distribution of $\Delta'$.

\begin{lemma}\label{lem:gaussian}
If the minibatch data is chosen randomly, then the distribution of $\Delta'$ is Gaussian.
\end{lemma}

\begin{proof}
By expanding the definition of $\Delta'$ from Equation~\ref{eq:deltas}, we get
\[
\Delta' = \log p(\theta') - \log p(\theta_t) + \log q(\theta_t \mid \theta') - \log q(\theta' \mid \theta_t) +
\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid \theta_t)).
\]
If we fix $\theta_t$ and $\theta'$, the randomness in our $\Delta'$ samples only comes from the
minibatch. Since the minibatch is chosen randomly, the distribution of $\sum_{i=1}^n (\log p(x_i\mid
\theta') - \log p(x_i\mid \theta_t))$ converge to a Gaussian by the Central Limit Theorem.
\end{proof}

From Lemma~\ref{lem:gaussian}, since $\Delta'$ is a noisy approximation of $\Delta$, the
relationship can be expressed as
\begin{equation}\label{eq:relationship}
\Delta' = \Delta + \epsilon, \quad \epsilon \sim \mathcal{N}(0, \sigma^2),
\end{equation}
for a Gaussian noise term $\epsilon$.  The extra noise in $\Delta'$ means we can no longer perform
the $\Delta + X$ test, so we need to slightly change our acceptance criteria. We can decompose
$X$ as $X = X_{\rm norm}+X_{\rm corr}$, where the first term is a zero-mean ``noise term'' which
has the same variance as $\epsilon$, and the second term is a zero-mean ``correction'' term.
These two convolve to form $X$. The criteria to accept can be rewritten as
\begin{equation}\label{eq:criteria}
\Delta + X = \Delta + X_{\rm norm} + X_{\rm corr} \approx \Delta' + X_{\rm corr} >0.
\end{equation}
This deconvolution process is visualized in Figure~\ref{fig:part2}. {\color{blue} (Daniel: note that
we will be fixing this figure, so the visualization should hopefully be a little clearer.)} We
decompose $X$ this way so that we can estimate the two components separately.  Notice that if we
increase the mini-batch size, this will cause the $X_{\rm noise}$ distribution to shrink and become
narrower.

Furthermore, since our proposed acceptance function $g(\Delta) = (1+ \exp(-\Delta))^{-1}$ satisfies
detailed balance (see Lemma~\ref{lem:detailed_balance}), if $\Delta$ is computed from the complete
data, the distribution of the sampled data will match our target distribution $f$, the posterior,
under standard MCMC assumptions (e.g., that the chain is run sufficiently long enough and mixes
well). Now consider the case of using minibatches of data. Let our true and estimated acceptance
probabilities be $P_a = g(\Delta) = \Pr(\Delta + X > 0)$ and $P_a' = \Pr(\Delta' + X_{\rm corr} >
0)$.  If our estimate of $X_{\rm norm}$ is exact, then $P_a' = P_a$, which follows by making the
inequality in Equation~\ref{eq:criteria} an equality.

Algorithm~\ref{alg:our_algorithm} concisely describes our MH test.

{\color{blue}
Daniel: I really need to make this clear and do the algorithm description.
}

\begin{comment}
though it really should be an approximation as we can't guarantee that our $X_{\rm norm}$ term is
exactly the correct amount of noise), where the left hand side has distributions representing normal
noise ($X_{\rm norm}$, which is labeled as ${\rm Noise}_U$, but they are the same thing) and a
``correction'' term labeled $X_{\rm correction}$. Be aware that all of these ``additions'' are to be
interpreted as convolutions when computing the sum of densities.

I think it's becoming clear but I'm still trying to connect all of the steps.  During each iteration
of MCMC, we compute $\Delta'$ and \emph{sample} (right?) an $X$ (which we can do by sampling $u$ and
then doing $X = g^{-1}(u)$. This gives us $\Delta' + X$ but our MH test actually requires $X_{\rm
correction}$, so first we must estimate $\sigma^2$, the variance of $X_{\rm noise} \sim
\mathcal{N}(0, \sigma^2)$ (perhaps from the mini-batch data, somehow?) and then somehow ``remove''
that component from $X$ to get $X_{\rm correction}$? Then we have our acceptance test. Is this
right? Or, of course, we don't have to sample $X$ at all, but just figure out a way to sample
$X_{\rm correction}$ directly.

I think we talked about estimating $\sigma^2$. The variance $\sigma^2$ that we're concerned about is
a scalar value based on the sum of log probabilities of different samples, so one way we could
estimate the variance is by taking all samples of a mini-batch and seeing how much those individual
$\log p(x_i \mid \theta)$ vary. Or we could look at $k$ mini-batches of data which will give us
$\Delta_1, \ldots, \Delta_k$ and compute the empirical $\sigma^2$ (of course that defeats the
purpose of using mini-batches).

I agree with you that out MH test requires $X_{correction}$. My understanding
is that we can actually estimate $X$ from the whole data set, and the probability distribution of
$X_{norm}$ is what we proposed? So the determination of $X_{correction}$ is a deterministic
deconvolution problem.
\end{comment}

\begin{algorithm}[t]
\KwData{this text}
\KwResult{how to write algorithm with \LaTeX2e } initialization\;
\While{not at end of this document}{
    read current\;
    \eIf{understand}{
        go to next section\;
        current section becomes this one\;
    }{
        go back to the beginning of current section\;
    }
}
\caption{A description of our MH test.}
\label{alg:our_algorithm}
\end{algorithm}

@@

In this section, we explore how well our algorithm will converge to the posterior distribution.
Define our estimate of $X_{\rm norm}$ as $X_{\rm norm}'$, and let $X_{\xi} = X_{\rm norm} - X_{\rm
norm}'$, so that $\Delta' = \Delta + X_{\rm norm} + X_\xi$.  We also denote $F_X$ as the CDF of $X$.
Let the transition kernel of complete data at sampling step $i$ be $T_i$, and let the transition
kernel of minibatch data at sampling step $i$ is $T_i'$.  We show that the distance of $T_i$ and
$T_i'$ is proportional to our estimate of $X_{\rm norm}$. We also define the stationary distribution
obtained by $T_i$ is $\pi$.  We show the by decreasing the step size, the distribution of our
sampled data converges to $\pi$.

\begin{lemma}
If $|X_\xi| < \zeta$ and $\nabla F_X < k$, the total variance distance between $T_i$ and $T_i'$ is
bounded by $2\zeta k$.
\end{lemma}

\begin{proof}
We first prove that the difference in our acceptance rates is bounded: $|P_a - P_a'| \leq 2\zeta k$
for some small constant $\zeta > 0$. Given $X_\xi = \xi$ the distance $|P_a - P_a'|$ is:
\begin{align*}
|P_a - P_a'| &= |\Pr(\Delta + X >0) - \Pr(\Delta' + X_{\rm corr} > 0)|\\
& = | \Pr (\Delta + X > 0) - \Pr (\Delta + X_{\rm norm} + \xi + X_{\rm corr} >0) |    \\
&= |\Pr (\Delta + X > 0) -\Pr(\Delta + X + \xi >0)| = |F_X(-\Delta)- F_X(-\Delta - \xi)| \\
&= |\nabla F(-\Delta) \xi + o(\xi^2)| \leq  |\nabla F(-\Delta) \xi| +| o(\xi^2)| \leq 2|\nabla F(-\Delta)| |\xi| \leq 2k \xi,
\end{align*}
where the last line uses Taylor's Theorem ($o(\xi^2)$ represents higher-order terms that have
smaller absolute value) and the Triangle Inequality.  Since $|X_\xi| < \zeta$, $|P_a - P_a'| <
2k\zeta$.

Next, we show the transition kernels at each step are bounded if we use the same proposer, i.e.
$\|T_i - T'_i\|_1 \leq 2k\zeta$ with $\| \cdot \|_1$ the standard $L_1$ norm. We can write the
$T_i(\theta_t \mid \theta') = P_a(\theta_t, \theta') q(\theta'\mid \theta_t) +
(1-P_a(\theta_t,\theta')) \delta_D(\theta' - \theta_t)$, where $\delta_D$ is the Dirac delta
function. The distance between two proposal kernels is
\begin{align*}
\|T_i - T'_i \|_1 &= \int_{\theta'} \left| \int_{\theta_t}((P_a-P_a') q(\theta'|\theta_t) + (1-P_a - 1+P_a') \delta_D(\theta_t -\theta')) d\theta_t  \right| d\theta' \\
& = \int_{\theta'} \left| \int_{\theta_t} (P_a- P_a')(q(\theta'|\theta_t) - \delta_D(\theta_t - \theta')) d\theta_t \right|  d\theta'\\ 
& \leq 2 \zeta k \int_{\theta'} \left| \int_{\theta_t}(q(\theta'|\theta_t) + q(\theta_t|\theta'))  d\theta_t \right| d\theta'= 2 \zeta k,
\end{align*}
as desired.
\end{proof}

Since the accuracy of our transition kernel is related to the accuracy of our $X_{\rm norm}$
estimate, we can show that error of our estimation is roughly proportional to the step size of our
proposer.

\begin{lemma}
For one step sampling from $\theta_t$ to $\theta'$, if the jumping step  $|\theta_t - \theta'| <
\epsilon$, and the gradient of the log-likelihood is bounded by a constant factor, i.e. $|\nabla
(\log \Pr(x_i| \theta))| < k$, the variance of $\Delta'$ is bounded by $4\epsilon^2 k^2 m$, where
$m$ is the minibatch size.
\end{lemma}

\begin{proof}
From Lemma~\ref{lem:gaussian}, only the $\sum_{i=1}^n (\log p(x_i\mid \theta') - \log p(x_i\mid
\theta_t))$ term brings randomness into $\Delta'$. To simplify the subsequent notation, we define
$g_i(\theta) = \log p(x_i\mid \theta)$. Then by Taylor's Theorem:
\[
|g(\theta') - g(\theta_t)| = |\nabla g(\theta_t) (\theta' - \theta_t) + o(\theta' - \theta_t)^2| \leq 2\epsilon k.
\]
Thus, with given the $\theta$, and $\theta'$, the variance of $\Delta'$ can be write as:
\begin{align*}
{\rm Var}(\Delta') &= {\rm Var}\left[\sum_{i=1}^n (\log \Pr(x_i | \theta') - \log \Pr (x_i |\theta_t))\right] = {\rm Var}\left[\sum_{i=1}^n g_i(\theta') - g_i(\theta_t)\right] \\
& \leq \sum_{i=1}^n {\rm Var}(g_i(\theta') - g_i(\theta_t)) \leq  4\epsilon^2 m k^2,
\end{align*}
as desired.
\end{proof}

Since ${\rm Var}(\Delta') < \epsilon^2 k^2 m$, we can assume the maximum value of $X_{\xi}$ is
proportional to the standard deviation of $\Delta$, i.e. $\epsilon$. Thus, we can assume
$\zeta=\epsilon C$, where $C$ is a constant factor.

\begin{theorem}
Assume the stationary distribution $\pi$ satisfies $\|T \circ P_0 - \pi\| \leq \eta \|P_0 - \pi\|$,
where $\eta \in [0, 1)$ for all distribution $P_0$. And assume there exists $\rho_t < 1$ such that
$\|T'_i(x, \cdot) - \pi_i'\| \leq 2\rho_i$. We can get $\| T_t' \circ T_{t-1}' \circ \cdots \circ T_1' P_0
- \pi_0 \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s \alpha_s + \alpha_t $,
where $\alpha_i = \frac{\epsilon_i k }{1-\eta}$.
\end{theorem}

\begin{proof}
First, by Theorem 1 in~\cite{cutting_mh_2014}, $\|T \circ P_0 - \pi\| \leq \eta \|P_0
- \pi\|$ and $\|T-T'_i\| \leq \epsilon_i k$, the norm between the stationary distribution is $\|\pi
- \pi_i'\|\leq \frac{\epsilon_i k}{1-\eta}$.

Second, by Theorem 3.6 in~\cite{yang2013sequential}, we have  $\| T_t' \circ T_{t-1}' \circ \cdots
\circ T_1' P_0 - \pi_t \| \leq \sum_{s=1}^t \{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\} \rho_s
\alpha_s$, where $\alpha = \frac{\epsilon_i k}{1-\eta}$. Thus, we get:
\begin{align*}
 \| T_t' \circ \cdots \circ T_1' P_0 - \pi_0 \| &\leq \|T_t' \circ \cdots \circ T_1' P_0 - \pi_t\| + \|\pi_t - \pi_0\| \\
 &\leq \sum_{s=1}^t \left\{\prod _{u=s+1}^t \rho_u (1-\alpha_u)\right\} \rho_s \alpha_s + \alpha_t,
\end{align*}
as desired.
\end{proof}
